---
title: "Long_Term_Assignment_Galih Trias Farobi"
author: "Galih Trias Farobi"
date: "10/7/2021"
output: html_document
---
First, step we have to prepare the Packages and library
```{r}
#Import Packages
pkgs <- c("missForest", "Hmisc", "VIM", "VGAM",
          "tidyverse", "reshape2", 'ggstatsplot', "cleandata", "caret",
          "forecast")

for (p in pkgs){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p,character.only = T)
}

install.packages('caTools')
install.packages("car")
install.packages("psych")
install.packages("car")
install.packages("glmnet")
install.packages("esquisse")
install.packages('ggplot2')
install.packages('dplyr')
install.packages('tidyr')
install.packages('patchwork')
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(tidyr)
library(patchwork)
library(tidyverse)
library(lubridate)
library(dplyr)
library(GGally)
library(e1071) 
library(rpart) 
library(rpart.plot) 
library(caTools)
library(psych)
library(car)
library(glmnet)
```
Then, we import the dataframe that we need. In this paper, the dataframe that we want to use is 'covid_19_indonesia_time_series_all'. This Dataframe was collected from January 8, 2020 to July 9, 2021. It contains 16,283 rows and 41 column.
```{r}
#Import The Dataframe
df <-  read_csv("C:/Users/User/Documents/R/covid_19_indonesia_time_series_all.csv")
df %>% view()
```

Next step, before we want to manipulating and analyze the dataframe that we have. We have to cleaning the data first. To prevent mismanipupaltion and misanalyze towards our data.
```{r}
#Data Cleaning
drop_cols <- c('Name',
               'Item',
               'Kind',
               'Hidden',
               'City or Regency',
               'Special Status',
               'Location ISO Code',
               'Province',
               'Country',
               'Continent',
               'Total Regencies',
               'Total Cities',
               'Total Districts',
               'Total Cases per Million',
               'New Deaths per Million',
               'Total Deaths per Million',
               'Case Fatality Rate',
               'Case Recovered Rate',
               'Growth Factor of New Cases',
               'Growth Factor of New Deaths',
               'Time Zone',
               'Total Urban Villages',  
               'Total Rural Villages',
               'New Cases per Million'
               )

df <- df %>% select(-drop_cols)
df %>% view() %>% str()
```
First step on Data cleaning on this Dataframe is to drop columns that actually can cause us ambiguity, and actually doesn't need it in
our process forward.

Next, is to replace values that actually ambiguous likes in column 'Location' and 'Island', in column 'Location' there is 'Indonesia' value, which actually pretty ambiguous in terms of name but the value that comes with it can still be taken into account, so because of that case we try to rename the 'Indonesia' value to 'Unknown'. Then, in column 'Island' there is NA value within it. Because the NA values within 'Island' it can be mean unknown island, so we decide to fill the NA values to 'Unknown'. In summary, because those two columns and the values that comes with themn can still be considered, so we decide not to drop the rows 'Indonesia' or NA, but we change it to Unknown so we still can process it.
```{r}
df$Location<-replace(df$Location, df$Location=='Indonesia','Unknown') 
df$Island <- ifelse(is.na(df$Island), 
             'Unknown', df$Island)
df %>% view()
```

Then, we check is there any NA value left in our dataframe
```{r}
indx <- apply(df, 2, function(x) any(is.na(x) | is.infinite(x)))
indx
```
After we check it, the result is no NA value anymore.

Now, we check is there any duplicates data in our dataframe
```{r}
duplicated(df)
sum(duplicated(df))
#There is no duplicates data in our dataframe
```
The result is there is no duplicates data in our dataframe. From here on, our data right now is ready to use for manipulating, because it is clean now.

```{r}
esquisse::esquisser()
```



Next process is EDA:
The pandemic has been going on for more than 17 months in Indonesia and until now (9 July 2021) it has not found a bright spot when the pandemic will end. If we look at the current data, will the pandemic be over as soon as possible? This data will give an idea of whether the pandemic that we have been experiencing can be controlled or on the contrary, which is getting worse.

```{r}
#Turn Date String Column to format Date
df$Date <- as.Date(df$Date, format = "%m/%d/%Y")
df
```
```{r}
#Separating Date with year, month, day
df$year <- year(ymd(df$Date))
df$month <- month(ymd(df$Date)) 
df$day <- day(ymd(df$Date))
df
```


First, we want to know how many total cases of Covid-19 in Indonesia until now (9 July 2021) for each Location.
```{r, fig.width=15}
df %>% 
  filter(!str_detect(Location, "Unknown")) %>% 
  group_by(Location) %>% 
  summarise(total_cases = sum(`Total Cases`)) %>% 
  arrange(desc(total_cases)) %>% 
  ggplot(aes(x=reorder(Location, total_cases), y=total_cases, fill = Location, weight = total_cases)) +
    geom_bar(stat="identity", alpha=.6, width=.4) +
    scale_fill_hue(direction = 1) + 
    scale_y_continuous(labels = function(x) format(x, scientific = FALSE))+
    coord_flip() +
    labs(
    x = "Location",
    y = "Total_Cases",
    title = "Total Cases in Indonesia per Location",
    subtitle = "Based on Covid-19 Indonesia Dataframe"
  ) +
    theme_bw()
```

it turns out Jakarta has the biggest total cases compare to any other Location, perhaps it's because Jakarta is close to the location of the first COVID-19 case, that is Depok. And Jakarta is the city with the highest mobilization in the JABODETABEK area.

Then we want to know in fact the most cases of covid 19 to date are on what island?

```{r}
df_island <- df %>% 
  filter(!str_detect(Island, "Unknown")) %>% 
  group_by(Island) %>% 
  summarise(total_cases_island = sum(`Total Cases`)) %>% 
  arrange(desc(total_cases_island))
df_island
```

```{r}
df_island %>% 
    ggplot(aes(x=reorder(Island, total_cases_island), y=total_cases_island, fill = Island, weight = total_cases_island)) +
    geom_bar(stat="identity", alpha=.6, width=.4) +
    scale_fill_hue(direction = 1) + 
    scale_y_continuous(labels = function(x) format(x, scientific = FALSE))+
    coord_flip() +
    labs(
    x = "Island",
    y = "Total Cases",
    title = "Total Cases in Indonesia per Island",
    subtitle = "Based on Covid-19 Indonesia Dataframe"
  ) +
    theme_bw()
```

it turns out 'Jawa' is dominating for total cases per Island in Indonesia. Also, 'Jawa' is the biggest total cases compare to any other Island in Indonesia. Perhaps because the first case of Covid-19 in Indonesia start in 'Jawa'.

Next, we want to know how the case progresses for each Location, in the 4 highest cases locations in Indonesia, we start from Jakarta that has highest case.
```{r}
Jakarta_LP <- df %>% 
  filter(!str_detect(Location, "Unknown")) %>% 
  filter(Location == 'DKI Jakarta') %>% 
  group_by(year, month, Date) %>% 
  summarise(total_New_Cases = sum(`New Cases`))

Jakarta_LP %>% 
  ggplot(aes(x=Date, y=total_New_Cases)) +
    geom_line( color="maroon") +
    theme_bw() +
    labs(
    x = "Date",
    y = "Total New Cases",
    title = "Growth of Total New Cases of Covid-19 in Jakarta",
    subtitle = "Based on Covid-19 Indonesia Dataframe"
  ) + 
  scale_x_date(date_breaks = '3 months', date_labels = "%B\n%Y")
```
As we can see in this paragraph that Covid-19 total new cases in Jakarta until now (9 july 2021),in 2020 the graphics tend to be stable but towards the end of the year new Covid-19 cases began to increase, maybe this happened as a result of a year-end event that occurred at the end of 2020 to early 2021, covid cases began to slowly decline again until the end of May period, but after that the total number of Covid cases increased dramatically to an unprecedented level in mid-2021. This is likely due to the Eid period that occurred in mid-2021

How about another Location? Next we need to see what happen in another location, next we will see'Jawa Barat'.
```{r}
JawaBarat_LP <- df %>% 
  filter(!str_detect(Location, "Unknown")) %>% 
  filter(Location == 'Jawa Barat') %>% 
  group_by(year, month, Date) %>% 
  summarise(total_New_Cases = sum(`New Cases`)) 

JawaBarat_LP %>% 
    ggplot(aes(x=Date, y=total_New_Cases)) +
    geom_line( color="green") +
    theme_bw() +
    labs(
    x = "Date",
    y = "Total New Cases",
    title = "Growth of Total New Cases of Covid-19 in West Java",
    subtitle = "Based on Covid-19 Indonesia Dataframe"
  ) +
  scale_x_date(date_breaks = '3 months', date_labels = "%B\n%Y")
```
Not much different from the data from 'Jakarta', at the beginning of the 2020 period, the total number of new cases of Covid-19 in 'Jawa Barat' was low and quite stable until the end of the year period, precisely at the end of 2020 to early 2021. However, what makes it interesting was the case in 'Jawa Barat' especially at the end of the year period. it jumped drastically at the end of the year. Maybe, it seems that there are many people who vacation in 'Jawa Barat' because of the year-end holidays. Then, after that, Covid-19 cases in 'Jawa Barat' gradually decreased, but by the time we entered the middle of the year, it turned out that 'Jawa Barat' had experienced a very drastic increase, exceeding the initial period of the year, maybe this was triggered by the Eid holiday where 'Jawa Barat' became a favorite destination.

Next, we will see the cases in 'Jawa Tengah'.
```{r}
JawaTengah_LP <- df %>% 
  filter(!str_detect(Location, "Unknown")) %>% 
  filter(Location == 'Jawa Tengah') %>% 
  group_by(year, month, Date) %>% 
  summarise(total_New_Cases = sum(`New Cases`))

JawaTengah_LP  %>% 
  ggplot(aes(x=Date, y=total_New_Cases)) +
    geom_line( color="darkblue") +
    theme_bw() +
    labs(
    x = "Date",
    y = "Total New Cases",
    title = "Growth of Total New Cases of Covid-19 in Central Java",
    subtitle = "Based on Covid-19 Indonesia Dataframe"
  ) +
  scale_x_date(date_breaks = '3 months', date_labels = "%B\n%Y")
```
not much different from the previous two cities, the Central Java area also experienced an increase at the end of the year, and began to be under control after the year-end holiday period. However, what is interesting here is that the total number of new cases in Central Java had decreased significantly in the mid-year period, to be precise during the Eid holiday, before it increased again significantly in the next periode

Then, we will new cases in East Java.
```{r}
JawaTimur_LP <- df %>% 
  filter(!str_detect(Location, "Unknown")) %>% 
  filter(Location == 'Jawa Timur') %>% 
  group_by(year, month, Date) %>% 
  summarise(total_New_Cases = sum(`New Cases`)) 

JawaTimur_LP %>% 
  ggplot(aes(x=Date, y=total_New_Cases)) +
    geom_line( color="darkgoldenrod2") +
    theme_bw() +
    labs(
    x = "Date",
    y = "Total New Cases",
    title = "Growth of Total New Cases of Covid-19 in East Java",
    subtitle = "Based on Covid-19 Indonesia Dataframe"
  ) + 
  scale_x_date(date_breaks = '3 months', date_labels = "%B\n%Y")
```
In East Java, the case is still the same, which is not too much different from what was in the previous three cities, but perhaps what needs to be highlighted is the end of the year period in a specifically close range period the case was very high, but then it suddenly dropped once before slowly rising again. After that, entering the post-year-end period, it began to stabilize before it exploded and never experienced a decline in the middle of the year, which coincided with the Eid holiday.

```{r, fig.width=15}
Java_LP <- ggplot() + 
  geom_line(data = Jakarta_LP, aes(x = Date, y=total_New_Cases), color = "maroon", size = 1.3) + #DKI Jakarta
  geom_line(data = JawaBarat_LP, aes(x = Date, y=total_New_Cases), color = "green", size = 1.3) + #Jawa Barat
  geom_line(data = JawaTengah_LP, aes(x = Date, y=total_New_Cases), color = "darkblue", size = 1.3) + #Jawa Tengah
  geom_line(data = JawaTimur_LP, aes(x = Date, y=total_New_Cases), color = "darkgoldenrod2", size = 1.3) + #Jawa Timur
  
  theme_gray() +
    labs(
    x = "Date",
    y = "Total New Cases",
    title = "Growth of Total New Cases of Covid-19 in Java",
    subtitle = "Based on Covid-19 Indonesia Dataframe"
  ) +
  scale_x_date(date_breaks = '3 months', date_labels = "%B\n%Y")

Java_LP
```
Maroon = 'DKI Jakarta'
Green = 'Jawa Barat'
Dark Blue = 'Jawa Tengah'
Gold = 'Jawa Timur'

if we look here that the graphic of the 4 largest locations of total new cases of Covid-19 in the Java area, it shows that all 4 locations experienced a "peak" period that was almost the same, namely at the beginning of the year in the period of December and January, then the largest was in June. -July. However, it turns out that if we look here, something interesting is that it turns out that the total number of new cases in West Java is greater than Jakarta at the turn of the year, and entering the Eid holiday, to be precise in the April-May period, the total number of cases in West Java is the highest in Indonesia. Java Island

```{r}
Java_df <- df %>% 
  filter(!str_detect(Island, "Unknown")) %>% 
  filter(Island == 'Jawa') %>% 
  group_by(year, month, Date, Island) %>%
  summarise(total_New_Cases = sum(`New Cases`)) %>%  
  arrange(desc(total_New_Cases))
  
Sumatra_df <- df %>% 
  filter(!str_detect(Island, "Unknown")) %>% 
  filter(Island == 'Sumatera') %>% 
  group_by(year, month, Date, Island) %>%
  summarise(total_New_Cases = sum(`New Cases`)) %>%  
  arrange(desc(total_New_Cases))
  
Borneo_df <- df %>% 
  filter(!str_detect(Island, "Unknown")) %>% 
  filter(Island == 'Kalimantan') %>% 
  group_by(year, month, Date, Island) %>%
  summarise(total_New_Cases = sum(`New Cases`)) %>%  
  arrange(desc(total_New_Cases))
```
```{r, fig.width=10}
Big_Island_LP <- ggplot() + 
  geom_line(data = Java_df, aes(x = Date, y=total_New_Cases), color = "maroon", size = 1.3) + #Jawa
  geom_line(data = Sumatra_df, aes(x = Date, y=total_New_Cases), color = "green", size = 1.3) + #Sumatera
  geom_line(data = Borneo_df, aes(x = Date, y=total_New_Cases), color = "darkblue", size = 1.3) + #Kalimantan
  
  theme_gray() +
    labs(
    x = "Date",
    y = "Total New Cases",
    title = "Growth of Total New Cases of Covid-19 in 3 Big Islands",
    subtitle = "Based on Covid-19 Indonesia Dataframe"
  ) +
  scale_x_date(date_breaks = '3 months', date_labels = "%B\n%Y")

Big_Island_LP
```
Maroon = Jawa
Green = Sumatera
Blue = Kalimantan

If we see here that the largest Covid-19 case in the initial period until now (9 July 2021), the number of Covid-19 cases in Java is indeed the largest. However, what is interesting here is that when looking at the development of cases from the beginning to the present, it turns out that both Sumatra and Kalimantan did not experience a 'peak' period at the beginning of the year, namely in the period from December to January. However, during the Lebaran holiday period, cases in these three islands have increased (for Java, this has increased significantly).

Next, how about we compare between new cases, new recover and new death. We will use 'DKI Jakarta' as an example. Because 'DKI Jakarta'is more representative of how the development of Covid-19 in Indonesia, both in terms of Location and Island.
```{r}
Jakarta_NC <- df %>% 
  filter(!str_detect(Location, "Unknown")) %>% 
  filter(Location == 'DKI Jakarta') %>% 
  group_by(year, month, Date) %>% 
  summarise(total_New_Cases = sum(`New Cases`))

Jakarta_NR <- df %>% 
  filter(!str_detect(Location, "Unknown")) %>% 
  filter(Location == 'DKI Jakarta') %>% 
  group_by(year, month, Date) %>% 
  summarise(total_New_Recover = sum(`New Recovered`))
  
Jakarta_ND <- df %>% 
  filter(!str_detect(Location, "Unknown")) %>% 
  filter(Location == 'DKI Jakarta') %>% 
  group_by(year, month, Date) %>% 
  summarise(total_New_Deaths = sum(`New Deaths`))
```
```{r}
Jakarta_NC_LP <- Jakarta_NC %>% 
  ggplot(aes(x=Date, y=total_New_Cases)) +
    geom_line( color="maroon", size = 1) +
    theme_bw() +
    labs(
    x = "Date",
    y = "Total New Cases",
    title = "New Cases"
  ) + 
  scale_x_date(date_breaks = '3 months', date_labels = "%B\n%Y") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

Jakarta_NR_LP <- Jakarta_NR %>% 
  ggplot(aes(x=Date, y=total_New_Recover)) +
    geom_line( color="darkblue", size =1) +
    theme_bw() +
    labs(
    x = "Date",
    y = "Total New Recover",
    title = "New Recover"
  ) + 
  scale_x_date(date_breaks = '3 months', date_labels = "%B\n%Y") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

Jakarta_ND_LP <- Jakarta_ND %>% 
  ggplot(aes(x=Date, y=total_New_Deaths)) +
    geom_line( color="darkgoldenrod2", size = 1) +
    theme_bw() +
    labs(
    x = "Date",
    y = "Total New Deaths",
    title = "New Deaths"
    ) + 
  scale_x_date(date_breaks = '3 months', date_labels = "%B\n%Y") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```


```{r,}
Jakarta_Case <- (Jakarta_NC_LP | Jakarta_NR_LP |Jakarta_ND_LP)
Jakarta_Case +
  plot_annotation(title = 'Jakarta Case of Covid-19', subtitle = 'Based on Covid-19 Indonesia Dataframe', caption = 'In terms of New Cases, New Recovery, and New Deaths data')
```
Maroon = New Cases
Blue = New Recover
Gold = Ne Death

As we can see based on the New Cases, New Recover, and New Death charts. This illustrates that there is a straight line between the increase in new cases and the increase in recoveries and deaths. However, if we look at the number of deaths until now (9 July 2021) compared to the number of recovered cases. The number of Recover is always greater than the number of deaths

If we see it toes-to-toes between those graph:
```{r}
NC_NR_ND_LP <- ggplot() + 
  geom_line(data = Jakarta_NC, aes(x = Date, y=total_New_Cases), color = "maroon", size = 1.3) + #New Cases
  geom_line(data = Jakarta_NR, aes(x = Date, y=total_New_Recover), color = "darkblue", size = 1.3) + #New Recover
  geom_line(data = Jakarta_ND, aes(x = Date, y=total_New_Deaths), color = "darkgoldenrod2", size = 1.3) + #New Deaths
  
  theme_gray() +
    labs(
    x = "Date",
    y = "Total Cases",
    title = "New Cases vs New Recover vs New Deaths of Covid-19 in Jakarta",
    subtitle = "Based on Covid-19 Indonesia Dataframe",
    caption = "In terms of New Cases, New Recovery, and New Deaths data"
  ) +
  scale_x_date(date_breaks = '3 months', date_labels = "%B\n%Y")

NC_NR_ND_LP
```
Maroon = New Cases
Blue = New Recover
Gold = Ne Death

If we compare all three together. It can be seen that, the total increase in new cases and the total number of recoveries in the period before the Eid holiday, the two graphs tend to go hand in hand, however, when the number of cases during the Eid holiday rises greatly, the number of recoveries cannot coincide but towards the end of this dataframe (i.e. on 9 July 2021), the recovery chart rose significantly.

But, we still can't determine whether is it because of government policy in "PPKM" so that cases in the post-year period have decreased or there is something. For this case we have to seperate the date when is date before and after PPKM actually applied
```{r}
df$PPKM = ifelse(df$Date >= "2021-01-11" & df$Date  <= "2021-07-02", "PPKM", 
                 ifelse(df$Date >= "2021-07-03","PPKM Darurat",
                        ifelse(df$Date >= "2020-04-17" & df$Date <= "2020-12-07","PSBB","No Policy")))
df
```

```{r}
PPKM_df <- df %>% 
  filter(PPKM == 'PPKM') %>% 
  group_by(year, Date) %>%
  summarise(total_New_Cases = sum(`New Cases`)) %>%  
  arrange(desc(total_New_Cases))

No_Policy_df <- df %>% 
  filter(PPKM == 'No Policy') %>% 
  group_by(year, Date) %>%
  summarise(total_New_Cases = sum(`New Cases`)) %>%  
  arrange(desc(total_New_Cases))

PPKM_Darurat_df <- df %>% 
  filter(PPKM == 'PPKM Darurat') %>% 
  group_by(year, Date) %>%
  summarise(total_New_Cases = sum(`New Cases`)) %>%  
  arrange(desc(total_New_Cases))

PSBB_df <- df %>% 
  filter(PPKM == 'PSBB') %>% 
  group_by(year, Date) %>%
  summarise(total_New_Cases = sum(`New Cases`)) %>%  
  arrange(desc(total_New_Cases))

PPKM_LP <- ggplot() + 
  geom_line(data = PPKM_df, aes(x = Date, y=total_New_Cases), color = "maroon", size = 1.3) + #PPKM
  geom_line(data = NOT_PPKM_df, aes(x = Date, y=total_New_Cases), color = "darkblue", size = 1.3) + #No Policy
  geom_line(data = PPKM_Darurat_df, aes(x = Date, y=total_New_Cases), color = "darkgoldenrod2", size = 1.3) + #PPKM Darurat
  geom_line(data = PSBB_df, aes(x = Date, y=total_New_Cases), color = "green", size = 1.3) +#PSBB
  
  theme_bw() +
    labs(
    x = "Date",
    y = "Total Cases",
    title = "PPKM vs PSBB vs PPKM Emergency of Covid-19 in Indonesia",
    subtitle = "Based on Covid-19 Indonesia Dataframe",
    caption = "In terms of PPKM, PSBB, PPKM Emergency data"
  ) +
  scale_x_date(date_breaks = '2 months', date_labels = "%B\n%Y") +
  theme(axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1))

PPKM_LP
```
Marron = PPKM
Green = PSBB
Blue = No Policy
Gold = PPKM Emergency

when the Indonesian government made a policy to applied PPKM which was implemented on January 3, 2021, it was seen that at the beginning of its implementation, PPKM managed to reduce cases due to the absence of policy of PSBB which was over at December 7, 2020, but the success of PPKM did not work due to many who chose to take a Eid holiday. Because the result of increase in new  cases at Eid Holiday, the Government implemented emergency PPKM (but in this dataframe it is not yet seen whether emergency PPKM has succeeded in reducing Covid-19 cases).

```{r}
PPKM_early_NC <- df %>% 
  filter(Date >= "2021-01-11" & Date <= "2021-01-18") %>% 
  group_by(year, month, Date) %>% 
  summarise(total_New_Cases = sum(`New Cases`))

PPKM_early_ND <- df %>% 
  filter(Date >= "2021-01-11" & Date <= "2021-01-18") %>% 
  group_by(year, month, Date) %>% 
  summarise(total_New_Deaths = sum(`New Deaths`))

PPKM_early_NR <- df %>% 
  filter(Date >= "2021-01-11" & Date <= "2021-01-18") %>% 
  group_by(year, month, Date) %>% 
  summarise(total_New_Recover = sum(`New Recovered`))

PPKM_early_LP <- ggplot() + 
  geom_line(data = PPKM_early_NC, aes(x = Date, y=total_New_Cases), color = "maroon", size = 1.3) + #New Cases
  geom_line(data = PPKM_early_NR, aes(x = Date, y=total_New_Recover), color = "darkblue", size = 1.3) + #New Recover
  geom_line(data = PPKM_early_ND, aes(x = Date, y=total_New_Deaths), color = "darkgoldenrod2", size = 1.3) + #New Deaths
  
  theme_gray() +
    labs(
    x = "Date",
    y = "Total Cases",
    title = "PPKM Early Implementation of Covid-19 in Indonesia",
    subtitle = "Based on Covid-19 Indonesia Dataframe",
    caption = "In terms of New Cases, New Recovery, and New Deaths data"
  ) +
  scale_x_date(date_breaks = '1 days', date_labels = "%d\n%B")

PPKM_early_LP
```
it can be seen from the graph above shows that in the first week of implementing PPKM in Indonesia, it turned out that from 5 days there was an increase in cases, but after 5 days the results were visible where it turned out that due to the implementation of PPKM new cases began to decline, and the cure rate finally increased , and the mortality rate tends to be stable

We can assume that the implementation of PSBB and PPKM in Indonesia has succeeded in reducing the number of cases, especially PPKM, and it seems that the Government of Indonesia should continue to implement a restrictive policy because it can effectively reduce the number of new cases in Indonesia (For Emergency PPKM, it is still not concluded because there is no sufficient data to determine whether the emergency PPKM was successful)

Prediction:
```{r}
str(df)
```
'New Cases','New Deaths','New Recovered','New Active Cases','Total Cases','Total Deaths','Total Recovered','Total Active Cases','Population','PPKM'
```{r}
drop_cols_Prediction <- c('Date',
                'Location',
               'Location Level',
               'Island',
               'Area (km2)',
               'Population Density',
               'Longitude',
               'Latitude',
               'year',
               'month',
               'day'
               )

new_df <- df %>% select(-drop_cols_Prediction)
new_df %>% view() %>% str()
```

```{r}
#one hot encoding for PPKM
library(caret)

dummy <- dummyVars(" ~ .", data=new_df)
new_df <- data.frame(predict(dummy, newdata = new_df))
new_df

colnames(new_df) <- c('New_Cases','New_Deaths','New_Recovered','New_Active_Cases','Total_Cases','Total_Deaths','Total_Recovered','Total_Active_Cases','Population','PPKM_No_Policy', 'PPKM_PPKM', 'PPKM_PPKM_Darurat', 'PPKM_PSBB')
new_df %>% str()
```

```{r}
#Prediction
#Split the data to Train, Validation, and Test data
set.seed(123)
sample <- sample.split(new_df$New_Deaths, SplitRatio = .80)
pre_train_df <- subset(new_df, sample == TRUE)
sample_train <- sample.split(pre_train_df$New_Deaths, SplitRatio = .80)

train_df <- subset(pre_train_df, sample_train == TRUE)
validation_df <- subset(pre_train_df, sample_train == FALSE)
test_df <- subset(new_df, sample == FALSE)
```

```{r, fig.width=10}
#Feature Selection
#Correlation study
library(psych)
pairs.panels(pre_train_df, 
             method = "pearson", # correlation method
             hist.col = "#00AFBB",
             density = TRUE,  # show density plots
             ellipses = TRUE # show correlation ellipses
)
```
New Cases - New Recovered 92% - New Death - drop New Recovered
New Cases - Total Cases 83% - New Death - drop Total Cases
New Cases - Total Death 82% - New Death - drop Total Death
New Cases - Total Recovered 81% - New Death - drop Total Recovered
New Cases - Total Active Cases 94% - New Death - drop Total Active Cases
New Recover - Total Cases 87%
New Recover - Total Death 85%
New Recover - Total Recover 85%
New Recover - Total Active Cases 95%
Total Cases - Total Death 98%
Total Cases - Total Active Cases 90%
Total Death - Total Recover 98%
Total Death - Total Active Cases 89%
Total Recover - Total Active Cases 88%

```{r}
# drop correlated columns
library(dplyr)
drop_cols_sample <- c('New_Recovered',
               'Total_Cases',
               'Total_Deaths',
               'Total_Recovered',
               'Total_Active_Cases')

train_df <- train_df %>% select(-drop_cols_sample)
validation_df <-  validation_df %>% select(-drop_cols_sample)
test_df <- test_df %>% select(-drop_cols_sample)
```

```{r}
# feature preprocessing
# to ensure we handle categorical features
x <- model.matrix(New_Deaths ~ ., train_df)[,-1]
y <- train_df$New_Deaths
```

```{r}
# ridge regression
# fit multiple ridge regression with different lambda
# lambda = [0.01, 0.1, 1, 10]
ridge_reg_pointzeroone <- glmnet(x, y, alpha = 0, lambda = 0.01)
coef(ridge_reg_pointzeroone)

ridge_reg_pointone <- glmnet(x, y, alpha = 0, lambda = 0.1)
coef(ridge_reg_pointone)

ridge_reg_one <- glmnet(x, y, alpha = 0, lambda = 1)
coef(ridge_reg_pointone)

ridge_reg_ten <- glmnet(x, y, alpha = 0, lambda = 10)
coef(ridge_reg_ten)
```
```{r}
# comparison on validation data
# to choose the best lambda

# Make predictions on the validation data
x_validation <- model.matrix(New_Deaths ~ ., validation_df)[,-1]
y_validation <- validation_df$New_Deaths

RMSE_ridge_pointzeroone <- sqrt(mean((y_validation - predict(ridge_reg_pointzeroone, x_validation))^2))
RMSE_ridge_pointzeroone #8.252002

RMSE_ridge_pointone <- sqrt(mean((y_validation - predict(ridge_reg_pointone, x_validation))^2))
RMSE_ridge_pointone #8.242678

RMSE_ridge_one <- sqrt(mean((y_validation - predict(ridge_reg_one, x_validation))^2))
RMSE_ridge_one #8.223498 <- best

RMSE_ridge_ten <- sqrt(mean((y_validation - predict(ridge_reg_ten, x_validation))^2))
RMSE_ridge_ten #8.682836
```
```{r}
# true evaluation on test data
# using the best model --> RMSE_ridge_pointzeroone
x_test <- model.matrix(New_Deaths ~ ., test_df)[,-1]
y_test <- test_df$New_Deaths

# RMSE
RMSE_ridge_best <- sqrt(mean((y_test - predict(ridge_reg_one, x_test))^2))
RMSE_ridge_best # 0.06423617

# MAE
MAE_ridge_best <- mean(abs(y_test-predict(ridge_reg_one, x_test)))
MAE_ridge_best

# MAPE
MAPE_ridge_best <- mean(abs((predict(ridge_reg_one, x_test) - y_test))/y_test) 
MAPE_ridge_best
```
For MAE it means that On average, our prediction deviates the true admit_prob by 3.3 and also for RMSE The standard deviation of
prediction errors is 8.7 ○ i.e. from the regression line, the residuals mostly deviate between +- 8.7, unfortunately we can define the MAPE because of there is zero value in this data

```{r}
############## LASSO
# lasso regression
# fit multiple lasso regression with different lambda
# lambda = [0.01, 0.1, 1, 10]
lasso_reg_pointzeroone <- glmnet(x, y, alpha = 1, lambda = 0.01)
coef(lasso_reg_pointzeroone) 

lasso_reg_pointone <- glmnet(x, y, alpha = 1, lambda = 0.1)
coef(lasso_reg_pointone) 

lasso_reg_one <- glmnet(x, y, alpha = 1, lambda = 1)
coef(lasso_reg_pointone)

lasso_reg_ten <- glmnet(x, y, alpha = 1, lambda = 10)
coef(lasso_reg_ten)
```
```{r}
# comparison on validation data
# to choose the best lambda
# Make predictions on the validation data
RMSE_lasso_pointzeroone <- sqrt(mean((y_validation - predict(lasso_reg_pointzeroone, x_validation))^2))
RMSE_lasso_pointzeroone # 8.249907 

RMSE_lasso_pointone <- sqrt(mean((y_validation - predict(lasso_reg_pointone, x_validation))^2))
RMSE_lasso_pointone # 8.223708 

RMSE_lasso_one <- sqrt(mean((y_validation - predict(lasso_reg_one, x_validation))^2))
RMSE_lasso_one # 8.153449 <- Best

RMSE_lasso_ten <- sqrt(mean((y_validation - predict(lasso_reg_ten, x_validation))^2))
RMSE_lasso_ten # 9.12958
```
```{r}
# true evaluation on test data
# using the best model --> RMSE_lasso_pointone
RMSE_lasso_best <- sqrt(mean((y_test - predict(lasso_reg_one, x_test))^2))
RMSE_lasso_best # 8.62048

# MAE
MAE_lasso_best <- mean(abs(y_test-predict(lasso_reg_one, x_test)))
MAE_lasso_best

# MAPE
MAPE_lasso_best <- mean(abs((predict(lasso_reg_one, x_test) - y_test))/y_test) 
MAPE_laso_best
```
For MAE it means that On average, our prediction deviates the true admit_prob by 3.3 and also for RMSE The standard deviation of prediction errors is 8.6 ○ i.e. from the regression line, the residuals mostly deviate between +- 8.6, unfortunately we can define the MAPE because of there is zero value in this data

Because There is zero value in the data that is why value of MAPE for both Ridge and Lasso respectively cannot be define.

Based on this model prediction with Regularization Regression, the model that we use for prediction of New Deaths is:
(Intercept)         -2.089367e-01
`New Cases`          2.262014e-02
`New Active Cases`  -7.382256e-03
Population           1.010345e-07
`PPKM_No Policy`    -3.248394e-01
PPKM_PPKM            2.416052e-01
`PPKM_PPKM Darurat`  7.187393e+00

New Deaths = -2.089367e-01 + (2.262014e-02 New Cases) + (-7.382256e-03 New Active Cases) +(1.010345e-07 Population) + (-3.248394e-01 No Policy) + (2.416052e-01 PPKM) + (7.187393e+00 PPKM Darurat)


